---
title: "Test example of bat weight analysis using _climwin_ package"
author: "Oleh Prylutskyi"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls()) # reset R's brain
path <- getwd()
# Vignettes
# library(climwin)
vignette("climwin", package = "climwin")
# vignette("advanced_climwin", package = "climwin")
```

## Data overview
In this example, we used measurements of forearm length (Ra) of bats as a response variable. We captured bats during summer mist-netting surveys in 7 years (Year) in two different locations (Territory). There also were oher variables we are going to test (e.g., body mass (W)).

Daily climate data were obtained using [_easiclimate_](https://cran.r-project.org/web/packages/easyclimate/index.html) package, and contain tree variables:

1. Prcp - Total precipitation in mm
2. Tmin - minimal temperature, Celsius degrees
3. Tmean - mean temperature, Celsius degrees

We will use only Tmin fo this example.

```{r load data}
# Load data
load(file = paste0(path, "/bat_data.Rdata"))
load(file = paste0(path, "/climate_data.Rdata"))
```

## Parameter setting
```{r}
# Load necessary libraries
library(climwin)
library(lme4)

# Set priors
cdate <- daily$date  # date var in climate data
bdate <- df$date     # date var in biological data

# Relative or absolute time response.
# If "absolute", reference day as day, month is needed
type <- "absolute"
refday <- c(01, 07)

# If "relative", time window calculated for each individual value
# type <- "relative"
# refday <- NULL

# Temporal resolution of climate data
# cinterval <- "day"
cinterval <- "week"

# Upper and lower limit for tested climate windows respectively
# Must correspond to the resolution chosen in ‘cinterval’
# if 'cinterval' set to "days"
# range <- c(250, 0)

# if 'cinterval' set to "week"
# range <- c(30, 0) # (~ 1 Dec)
# range <- c(23, 0) # ~ 15 Jan
range <- c(13, 0) # ~ 1 Apr

stat <- "mean"

func <- "lin"

# number of repeats for randomization
repeats <- 5 # for metric = "C"
# repeats <- 200 # for metric = "AIC"

# Metrics for assessing
metric = "C"
# metric = "AIC"

# Number of years in data
sample.size = nlevels(as.factor(as.vector(df$Year)))

# List of dependent variable candidates
xvar <- list(Tmin = daily$Tmin)
```

Define a baseline model.

We tried both the simplest one and one wit interaction  (sex), but both returned same issues.
```{r baseline}
# Null model with no climate signal
baseline <- lm(Ra ~ 1, data = df)

# # Null model with interaction
# # Firstly, it is necessary to create a variable called ‘climate’ in the biological dataset.
# df$climate <- 1
# # Set baseline model
# baseline <- lm(Ra ~ climate*sex, data = df)
```

## Candidate model fitting
```{r, message=FALSE, warning=FALSE}
# Fit candidate model set
RaWin <- slidingwin(xvar = xvar,
                    cdate = daily$date,
                    bdate = df$date,
                    baseline = baseline,
                    cinterval = cinterval,
                    range = range,
                    type = type, refday = refday,
                    stat = stat,
                    func = func,
                    # spatial = list(df$place_id, daily$place_id)
                    spatial = list(df$Territory, daily$Territory)
                    )

# Fit randomized model set for evaluation purposes
RaRand <- randwin(repeats = repeats,
                  xvar = xvar,
                  cdate = cdate,
                  bdate = bdate,
                  baseline = baseline,
                  cinterval = cinterval,
                  range = range,
                  type = type, refday = refday,
                  stat = stat,
                  func = func,
                  # spatial = list(df$place_id, daily$place_id)
                  spatial = list(df$Territory, daily$Territory)
                  )
```

## Explore the model(s)

```{r}
# Possible combinations of model parameters
RaWin$combos
```

Dataset preview

```{r}
head(RaWin[[1]]$Dataset)
```


Check out the best model

```{r}
RaWin[[1]]$BestModel
```

Check out the bets model's data

```{r}
head(RaWin[[1]]$BestModelData)
```

Check out the randomized model

```{r}
head(RaRand[[1]])
```

Plot residuals against fitted values to check dependency function (linear, quadratic, cubic, etc.)
```{r}
bestmod <- RaWin[[i]]$BestModel

# Create residuals vs fitted plot
ggplot(bestmod, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth() +
  labs(title = "Residuals vs. Fitted Plot for the best model candidate",
       x = "Fitted Values",
       y = "Residuals")
```

Estimate how likely our observed result would be at random

```{r}
pvalue(dataset = RaWin[[i]]$Dataset,
       datasetrand = RaRand[[i]],
       metric = metric,
       sample.size = sample.size)
```

```{r}
RaOutput <- RaWin[[1]]$Dataset
RaRand_data <- RaRand[[1]]
plothist(dataset = RaOutput, datasetrand = RaRand_data)
```

Plot the distribution of ΔAICc values across all tested climate windows.
```{r}
plotdelta(dataset = RaOutput)
```

Plot model weights
```{r}
plotweights(dataset = RaOutput)
```

Plot model coefficients
```{r}
plotbetas(dataset = RaOutput)
```

Median window
```{r}
plotwin(dataset = RaOutput)
```

Best model plot
```{r}
# Plot all diagnostic plots for a given parameter combination
WindowOpen <- RaWin[[1]]$Dataset[1, 2]
WindowClose <- RaWin[[1]]$Dataset[1, 3]

# Fit single best model
RaSingle <- singlewin(xvar = xvar,
                      cdate = cdate,
                      bdate = bdate,
                      baseline = baseline,
                      cinterval = cinterval,
                      range = c(WindowOpen, WindowClose),
                      type = type, refday = refday,
                      stat = stat,
                      func = func,
                      # spatial = list(df$place_id, daily$place_id)
                      spatial = list(df$Territory, daily$Territory)
                      )

plotbest(dataset = RaOutput,
         bestmodel = RaSingle$BestModel, 
         bestmodeldata = RaSingle$BestModelData)
```

Here you can observe the main issue: "Climate variable" axis, which I expected be shown actual temperatures, displays something strange. Since we are working with late spring-summer temperature in temperate Europe, there is no chance to have negative values here.

Substitute the output of climwin::singlewin() by the original best model (RaWin[[i]]$BestModel)

```{r}
RaSingle <- RaWin[[i]]$BestModel
plotbest(dataset = RaOutput,
         bestmodel = RaSingle$BestModel, 
         bestmodeldata = RaSingle$BestModelData)
```



```{r}
ggplotRegression <- function (fit) {

  require(ggplot2)
  
  ggplot(fit$model, aes_string(x = names(fit$model)[2], y = names(fit$model)[1])) + 
    geom_point(alpha = 0.7) +
    stat_smooth(method = "lm", col = "red") +
    labs(title = paste("Adj R2 = ",signif(summary(fit)$adj.r.squared, 5),
                       "Intercept =",signif(fit$coef[[1]],5 ),
                       " Slope =",signif(fit$coef[[2]], 5),
                       " P =",signif(summary(fit)$coef[2,4], 5)))
}
```



```{r}
ggplotRegression(bestmod) +
  labs(x = "Min Temp",
       y = "Ra")
```

```{r}
# Create residuals vs fitted plot
ggplot(bestmod, aes(x = .fitted, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = 2) +
  geom_smooth() +
  labs(title = "Residuals vs. Fitted Plot for the best model candidate",
       x = "Fitted Values",
       y = "Residuals")
```


```{r}
climwin::singlewin
```

```{r}
RaSingle$BestModelData
```





## Plot all diagnostics


```{r}
# png("climwin_Ra.png", width = 32 , height = 22, 
#     units = "cm", res = 300)
plotall(dataset = RaOutput,
        datasetrand = RaRand_data,
        bestmodel = RaSingle$BestModel,
        bestmodeldata = RaSingle$BestModelData)
# The issue returned, probably something with "ModelBeta" object/variable.
# dev.off()
```
