---
title: "Climate"
author: "Alona"
date: "2024-01-08"
output: html_document
editor_options: 
  markdown: 
    wrap: 100
---

# How weather conditions during pregnancy and lactation of bats influence subadult bats development?

# Material & Methods

## Bat Capture and Monitoring:

Bats were captured during July in the years 2008, 2011, 2014, and 2019 as part of a bat monitoring
program of the National Nature Park “Homilshansky lisy”, located in the Kharkiv Oblast, Ukraine. The
capture methodology involved the use of nylon mist nets (12 х 3 meters, mesh size - 15 mm). Mist
nets were consistently deployed at the same 9 locations each year of the monitoring. A total of 18
capture nights were conducted annually, with mist nets set from sunset until sunrise. Body weight
measurements, rounded to the nearest 0.1 gram, and forearm length measurements, rounded to the
nearest 0.1 millimeter, were obtained within 8-10 hours following the capture events.

Data collected during the capture process included bat species identification, sex and age
determination. Age classification was primarily based on the degree of ossification observed in the
joints of the finger phalanx of the wing, the size and morphology of nipples for female bats, and
for *Nyctalus noctula* specifically, the wear pattern of canine teeth. All captured bats underwent
biometric assessments before being promptly released back into their natural habitat. All procedures
involved in the capture, handling, and transportation of bats were conducted with strict adherence
to ethical standards and principles of animal welfare and conservation, as outlined by Gannon
(2007).

Weather Data Acquisition and Analysis: Weather data, encompassing precipitation (Prcp), minimum
temperature (Tmin), and maximum temperature (Tmax), for the study area were obtained using the
easyclimate R package. Additional weather variables, including diurnal temperature range, mean daily
temperature, positive daily temperature, and Growing Degree Units (the sum of mean daily
temperatures above 0°C from the commencement of the year), were computed for further analysis.

The analysis focused on two distinct periods deemed crucial for bat growth: the spring period,
spanning from March 25th to May 24th, characterized as the pre-natal phase for bat pups, and the
summer period, extending from May 25th to June 24th, considered as the early post-natal phase
preceding the captures. Due to the occurrence of a leap year in 2008, adjustments were made for
Julian days accordingly.

Research Question and Hypotheses: Whether weather conditions during the periods of pregnancy and
lactation influence the growth of young bats? It was hypothesized that years characterized by warmer
and drier conditions during spring would correspond to larger sizes in subadult bats.

A total of 736 subadult bats from two species, Nyctalus noctula and Myotis daubentonii were taken
for the analysis.

## Підготовка даних про довжину передпліччя Ra та вагу W кажанів

Data preparation

```{r, echo=FALSE, message=FALSE}
rm(list = ls()) # reset R's brain
library(lubridate)  
library(ggplot2)
library(tidyverse) # for data manipulations
library(easyclimate) # for daily climate data extraction
library(sf) # for spatial operations
# library(patchwork)
# library(jtools)

# Loaded packages with their versions
sessionInfo() # to check your locale

# if LC_TIME=uk_UA.UTF-8, your months will be in Ukrainian.
# To fix that, you need to re-map locate into US (will be back after restart R session)

# Set locale to get English months names
# Sys.setlocale("LC_TIME", "en_US.UTF-8") # for Linux
# Sys.setlocale("LC_TIME", "English") # for Windows

```

```{r}
df <- read.csv("Homolsha_all_years.csv") # read data from *.csv file

# Set variables classes
df$date <- as.Date(df$Data, "%d.%m.%Y") # convert date to R Data format
df$Place <- as.factor(df$Place)
df$Species <- as.factor(df$Species)
df$sex <- as.factor(df$sex)
df$age <- as.factor(df$age)

df$Year <- year(df$date) #виокремлення року
df$Year <-as.factor (df$Year)
#створення нової колонки з даними про день від початку року 
df$day <- yday(df$date)


df$Ra <- sub(",", ".", df$Ra) #заміна ком на крапки в промірах передпліччя
df$W <- sub(",", ".", df$W)

df$Ra <-as.numeric (df$Ra) #присвоєння промірам нумеричного типу даних
df$W <-as.numeric (df$W)

# Delete unused variables
df$Data <- NULL
df$reproductive.status <- NULL

# Take a look at the data
summary(df)

```

Check age structure among species

```{r}
table (df$Species,df$age)
```

Select only subadult Noctule bats

```{r}
# Subset only some species (e.g., N.noc. , M.daubentonii)
df <- subset(df,
              Species %in% c('Nyctalus noctula'
                              #,'Myotis daubentonii'
                             ) &
                age == "sad"
              )

summary (df)
```

Check how Ra changed with W
```{r}
ggplot(data = df, aes(Ra, W
                      # , colour = Species
                      )
       ) +
  geom_point() + 
  geom_smooth(method = "lm") + theme_classic()
```


## Make synthetic growth indices (index of body condition)

```{r}
# W/Ra
df$W_Ra <- df$W / df$Ra

# Ra / square root of W
df$Ra_sqW <- df$Ra / sqrt(df$W)

# Ra / log W
df$Ra_logW <- df$Ra / log(df$W)

# Ra / square root of day-of-year
df$Ra_sqDOY <- df$Ra / sqrt(df$day)

# log10 Ra
df$Ra_log10 <- log10(df$Ra)

# log10 W
df$W_log10 <- log10(df$W)

# Since day of year is not changing hugely during sampling time span in absolute 
# measure, scaling to July, 1st = "zero day" may be a possible alternative

# Ra / square root of day-since-Jul1
# Which day-of-year is July, 1st
Jul_1st <- tibble(
  year = c(2008, 2011, 2014, 2019),
  Jul1_DOY = c(lubridate::yday(as.Date("2008-07-01")),
               lubridate::yday(as.Date("2011-07-01")),
               lubridate::yday(as.Date("2014-07-01")),
               lubridate::yday(as.Date("2019-07-01"))
  )
)

Jul_1st
```

```{r}
# Only 2008 as a leap year, so we can calculate ratio with zero-day using
# the following loop:

for(i in 1: nrow(df)){
    if(df$Year[i] == "2008"){
    df$Ra_sqDsJ1[i] <- df$Ra[i] / (df$day[i] - 183)
    } else {
      df$Ra_sqDsJ1[i] <- df$Ra[i] / (df$day[i] - 182)
    }
}


```




```{r, message=FALSE, warning=FALSE}
# t ~ Ra
p_Ra_day_4years <- ggplot(data = df, aes(x = day, y = Ra, colour = Year)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  ggtitle("Ra ~ DOY") + 
  theme_classic()

# t ~ W
p_W_day_4years <- ggplot(data = df, aes(x = day, y = W, colour = Year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  ggtitle("W ~ DOY") + 
  theme_classic()

p_Ra_day_4years
p_W_day_4years

rm(p_Ra_day_4years, p_W_day_4years)
```


## Перевірка чи дані по Ra та W впродовж двох тижнів різняться суттєво



## fit a linear regression to Ra and W. Check if the Ra, W parametres changes with time aren't significant. So we can use it (for each year) as a one sample

```{r}
mod <- lm(Ra ~ day, data=subset(df,Year == '2008'))
summary (mod)                                            
plot(mod)
ggplot(data = subset(df,Year == '2008'), aes(x = day, y = Ra)) +
  geom_point() +
  geom_smooth(method = "lm") 

ggplot(data=subset(df,Year == '2008'), aes (x = Ra)) +
  geom_histogram()
```

## Check different growth metrics for variation during inventory time
Custom function to check dependency of selected growth metric on day across 4 years

```{r}
check_my_var <- function(data, dependent_var, independent_var) {
  
  formula <- formula(paste0(dependent_var, " ~ ", independent_var))

  years <- list("2008", "2011", "2014", "2019")
  
  yearly_stat <- list()
  
  for (i in 1:length(years)) {
    
    # Fit simple linear model
    mod <- lm(formula = formula, data = data[data$Year == years[i],])
    
    # Get response coefficient
    coef <- mod$coefficients[2]
    
    # Get p-value from model summary
    p_val <- summary(mod)$coefficients[,"Pr(>|t|)"][2]
    
    yearly_stat[[i]] <- c(coef, p_val)
    
  }
  
  yearly_stat <- as.data.frame(t(as.data.frame(yearly_stat)))
  rownames(yearly_stat) <- years
  colnames(yearly_stat) <- c("Coefficient", "p_value")
  
  # yearly_stat$sign_level <- ifelse(yearly_stat$p-value > 0.05, 
  #                                  " ", 
  #                                  ifelse(yearly_stat$p-value <= 0.05 & yearly_stat$p-value > 0.01,
  #                                         "*",
  #                                         "***")
  #                                  )
  for (k in 1:nrow(yearly_stat)) {
    
    yearly_stat$sign_level[k] <- if(yearly_stat$p_value[k] > 0.05) {" "} else {"*"}
    
  }
  
  p <- ggplot(data = data, aes_string(x = independent_var, y = dependent_var)) +
    geom_point() +
    geom_smooth(method = "lm") +
    facet_wrap(vars(Year)) +
    theme_classic()
  
  return(list(yearly_stat, p))
}
```

### Ra
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra", "day")
```


### W_Ra
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "W_Ra", "day")
```

### Ra_sqW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_sqW", "day")
```


### Ra_logW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_logW", "day")
```

### Ra_sqW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_sqW", "day")
```


### Ra_sqW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_sqW", "day")
```


### Ra_sqW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_sqW", "day")
```


### Ra_sqW
```{r, warning=FALSE, message=FALSE}
check_my_var(df, "Ra_sqW", "day")
```

# Test null models of random effect of years on bat growth

Instructions from <https://tobiasroth.github.io/BDAEcology/lmer.html> Additional beginner's guide
<https://rstudio-pubs-static.s3.amazonaws.com/63556_e35cc7e2dfb54a5bb551f3fa4b3ec4ae.html>

```{r, message=FALSE}
library (lme4)
# `lme4` requires `Matrix` >= 1.6-4. See for more details https://community.rstudio.com/t/error-in-initializeptr-function-cholmod-factor-ldeta-not-provided-by-package-matrix/178694
```

Null-model (random effect only)

```{r}
mod0Ra <- lmer (Ra ~ 1 + (1|Year),
             data=nnoc_df
             # REML=F # https://tobiasroth.github.io/BDAEcology/lmer.html#restricted-maximum-likelihood-estimation-reml
             )

summary (mod0Ra)
```

```{r}
summ(mod0Ra)
```

To test the variance of intersepts

```{r}
library(lmerTest)
ranova(mod0Ra)
```

Model with one fixed effect - time (as day from the beginning of year)

```{r}
mod1Ra <- lmer (Ra ~ day + (1 |Year),
             data=nnoc_df
             # REML=F # https://tobiasroth.github.io/BDAEcology/lmer.html#restricted-maximum-likelihood-estimation-reml
             )

summary (mod1Ra)
```

```{r}
library(jtools)
summ(mod1Ra)
```

Tp test the variance of intersepts

```{r}
library(lmerTest)
ranova(mod1Ra)
```

```{r}
plot (mod1Ra)
```

## t \~ W

Make a model with Weight gata of subadult bast Null-model (random effect only)

```{r}
mod0_W <- lmer (W ~ 1 + (1|Year),
             data=nnoc_df
             # REML=F # https://tobiasroth.github.io/BDAEcology/lmer.html#restricted-maximum-likelihood-estimation-reml
             )


summ(mod0_W)
```

Model weight with one fixed effect - time (as day from the beginning of year)

```{r}
mod1W <- lmer (W ~ day + (1 |Year),
             data=nnoc_df
             # REML=F # https://tobiasroth.github.io/BDAEcology/lmer.html#restricted-maximum-likelihood-estimation-reml
             )

summ(mod1W)
```

```{r}
library(lmerTest)
ranova(mod1W)
```

```{r}
plot (mod1W)
```

```{r}
ggplot(data = df, aes(x = day, y = W, colour = Year)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(vars(Species),
             scales = "free") 
```

## Model for body condition Ra-W

створення нової колонки з даними про співідношення Ra до W

```{r}
nnoc_df$Ra_W <- nnoc_df$Ra / nnoc_df$W
nnoc_df$Ra_sqrtW <- nnoc_df$Ra / sqrt (nnoc_df$W)
nnoc_df$Ra_logW <- nnoc_df$Ra / log (nnoc_df$W)
nnoc_df$Ra_sqrtday <- nnoc_df$Ra / sqrt (nnoc_df$day)
nnoc_df$Ra_logday <- nnoc_df$Ra / log (nnoc_df$day)
```

## Generate weather data

```{r}
# Take a look at the timespan
summary(df$date)
```

```{r eval=FALSE, echo=FALSE, message=FALSE}
# # Create data frame with point coordinates
# coords <- data.frame(
#   place = c("R1", "R2", "Fed1", "Fed2", "Fed3", "F1", "F2", "F3", "F4"),
#   lon = c(36.33329, 36.32946, 36.32732, 36.32863, 36.33436, 36.32396, 36.32319, 36.31990, 36.31909),
#   lat = c(49.62554, 49.62092, 49.62298, 49.61411, 49.60477, 49.61782, 49.61169, 49.60948, 49.60782)
# )
# 
# # Retrieve daily weather values using {easyclimate} package
# # Prcp - Total precipitation in mm
# # Tmin - minimal temperature, Celsius degrees
# # Tmax - maximal temperature, Celsius degrees
# 
# daily <- get_daily_climate(
#   coords = coords,
#   period = 2008:2019,
#   climatic_var = c("Prcp","Tmin","Tmax"),
#   output = "df")
# 
# save(daily, file = "daily_weather_2008-2019.Rdata") # save weather data
```

```{r echo=FALSE}
load(file = "daily_weather_2008-2019.Rdata")
```

### Calculate additional weather variables

Calculate derived metrics (diurnial range, mean daily temperature, positive daily temperature).
Subset only years of inventory (2008, 2011, 2014, 2019)

```{r}
daily <- daily %>% 
   # Calculate derived metrics for each day
  mutate(DiurnialRange = Tmax - Tmin,
         MeanTemp = (Tmin + Tmax) / 2,
         # Get only positive mean temperatures (for Growing degree units calculation)
         PositiveMeanTemp = ifelse(MeanTemp > 0, MeanTemp, 0)) %>%
  # Adjust dates and add new variables for month and years
  mutate(date = as.Date(date),
         month = months(date),
         year = format(date, format = "%Y")) %>%
  # Filter only years of inventories
  filter(year %in% c(2008, 2011, 2014, 2019))
```

### Calculate GDU

Growing degree units (Sum of the mean daily temperature above 0 from the start of the year). Create
custom function that calculate GDUs for each day of the year.

```{r}
# Must be calculated for each year separately.
# Data must contain one row per day and the variable with the positive daily mean temperature.
# If mean temperature is negative, put zero in this variable.
# Also the variable with the full date MUST be character AND named "date"
gdu <- function(daily_data_for_one_year, 
                var_with_PositiveMeanTemp) {
  GDU <- list() # prepare empty list
  
  # looping through the daily data within the year 
  for (i in 1:nrow(daily_data_for_one_year)) {
    GDU[[i]] <- c(daily_data_for_one_year$date[i],
                  sum(daily_data_for_one_year[1:i, var_with_PositiveMeanTemp])
    )
  }
  # reshape obtained list
  GDU <- as.data.frame(GDU)
  GDU <- as.data.frame(t(GDU))
  colnames(GDU) <- c("date", "GDU")
  # retrieve the result
  return(GDU)
}
```

Calculate GDU and add it to the main weather data frame.

```{r}
# List all required years
list_of_years <- c(2008, 2011, 2014, 2019)

# Create an empty list
gdu_by_years <- list()

for (k in 1:length(list_of_years)) {
  # Calculate GDUs for each year
  df_temp <- daily %>% 
    # loop doesn't seem to work with `date` format, convert to character
    mutate(date = as.character(date)) %>%
    filter(year == list_of_years[[k]]) %>% 
    select(place, date, PositiveMeanTemp) %>%
    split(.$place)
  
  # Create an empty list
  df_gdu <- list()
  
  # Loop through the list elements (data frames with all days for the particular year,
  # split by place ID)
  for (i in 1:length(df_temp)) {
    df_gdu[[i]] <- gdu(df_temp[[i]], "PositiveMeanTemp")
    df_gdu[[i]]$place <- df_temp[[i]]$place
  }
  
  # Transform list of data frames into a single data frame
  gdu_by_years[[k]] <- bind_rows(df_gdu)

}

# Transform list of data frames into a single data frame
GDU <- bind_rows(gdu_by_years) %>% 
  mutate(date = as.Date(date))

# Add GDU as additional variable to the daily climate data for required years
daily <- daily %>% 
  left_join(GDU, by = c("date", "place"))

# Drop temporary variables not longer required
rm(df_temp, df_gdu, GDU, gdu_by_years, i, k)
```

## Explore weather data

We selected two periods that are supposed to be influential for bar growth: the *spring* (March 25 -
May 24) as a pre-natal period for bat pups, and the *summer* (May 25 - June 24) as an early
post-natal period prior to the captures.

Since 2008 was a leap year, Julian days was different. What is March 25 for each year?

```{r}
daily %>% 
  mutate(day = lubridate::yday(date)) %>% 
  select(date, day) %>% 
  filter(month(date) == 3 & day(date) == 25) %>% 
  unique()
```

Excluding dates out of those two periods

```{r}
# Pre-natal (spring) periods only
daily %>% 
  filter(date >= "2008-03-25" & date < "2008-05-25") %>% 
  mutate(season = "spring") -> spring2008

daily %>% 
  filter(date >= "2011-03-25" & date < "2011-05-25") %>% 
  mutate(season = "spring") -> spring2011

daily %>% 
  filter(date >= "2014-03-25" & date < "2014-05-25") %>% 
  mutate(season = "spring") -> spring2014

daily %>% 
  filter(date >= "2019-03-25" & date < "2019-05-25") %>% 
  mutate(season = "spring") -> spring2019

# Post-natal (summer) periods only
daily %>% 
  filter(date >= "2008-05-25" & date < "2008-06-25") %>% 
  mutate(season = "summer") -> summer2008

daily %>% 
  filter(date >= "2011-05-25" & date < "2011-06-25") %>% 
  mutate(season = "summer") -> summer2011

daily %>% 
  filter(date >= "2014-05-25" & date < "2014-06-25") %>% 
  mutate(season = "summer") -> summer2014

daily %>% 
  filter(date >= "2019-05-25" & date < "2019-06-25") %>% 
  mutate(season = "summer") -> summer2019

daily_spring_summer <- bind_rows(spring2008, spring2011, spring2014, spring2019,
                                 summer2008, summer2011, summer2014, summer2019)

# Remove unused temorary variables
rm(spring2008, spring2011, spring2014, spring2019,
   summer2008, summer2011, summer2014, summer2019)
```

Take a look at the minimal temperatures for springs in 2008-2019

```{r}
daily_spring_summer %>% 
  filter(season == "spring") %>% 
  select(date, year, Tmin) |>
  ggplot(aes(Tmin)) +
  geom_density(fill = "lightgreen", 
               alpha = 0.5) +
  facet_wrap( ~ factor(year)) +
  labs(title = "Spring",
       x = "Minimal temperature") +
  theme_bw()

ggsave("./figures/Tmin_2008-2019_density_Spring.png", width = 16, height = 12, units = "cm", dpi = 150)
```

Take a look at the minimal temperatures for summers in 2008-2019

```{r}
daily_spring_summer |>
  filter(season == "summer") %>% 
  select(date, year, Tmin) |>
  ggplot(aes(Tmin)) +
  geom_density(fill = "lightgreen", 
               alpha = 0.5) +
  facet_wrap( ~ factor(year)) +
  labs(title = "Summer",
       x = "Minimal temperature") +
  theme_bw()

ggsave("./figures/Tmin_2008-2019_density_Apr.png", width = 16, height = 12, units = "cm", dpi = 150)
```

## Generate exploratory variables for bat growth \~ weather models

```{r}
# Cumulative precipitation
daily_spring_summer %>% 
  group_by(season, place, year) %>% 
  summarise(sum(Prcp)) %>% 
  rename(Prcp = 4) %>% 
  as_tibble() -> Prcp_var

# Split spring only precipitation
Prcp_var %>% 
  filter(season == "spring") %>% 
  rename(Prcp_spr = Prcp) %>% 
  select(-season) -> Prcp_spring

# Split summer only precipitation
Prcp_var %>% 
  filter(season == "summer") %>% 
  rename(Prcp_summ = Prcp) %>% 
  select(-season) -> Prcp_summer


# Average minimum daily temperature
daily_spring_summer %>% 
  group_by(season, place, year) %>% 
  summarise(mean(Tmin)) %>% 
  rename(Tmin = 4) %>% 
  as_tibble() -> Tmin_var

# Dissect only spring minimal temperatures
Tmin_var %>% 
  filter(season == "spring") %>% 
  rename(Tmin_spr = Tmin) %>% 
  select(-season) -> Tmin_spring

# Dissect only spring minimal temperatures
Tmin_var %>% 
  filter(season == "summer") %>% 
  rename(Tmin_summ = Tmin) %>% 
  select(-season) -> Tmin_summer

# Number of days with mean temperature less or equal 10 Celsius degree
daily_spring_summer %>% 
  filter(season == "spring") %>% 
  filter(MeanTemp <= 10) %>% 
  count(season, place, year) %>% 
  as_tibble() %>% 
  rename(Less10_var = n) %>% 
  select(-season) -> Less10_var


# GDU at the end of a spring period
daily_spring_summer %>% 
  filter(month(date) == 5 & day(date) == 24) %>% 
  select(place, year, GDU) %>% 
  rename(GDU_spr = GDU) -> GDU_spring

# GDU at the end of a summer period
daily_spring_summer %>% 
  filter(month(date) == 6 & day(date) == 24) %>% 
  select(place, year, GDU) %>% 
  rename(GDU_summ = GDU) -> GDU_summer

# Merge weather covariates
Prcp_spring %>% 
  left_join(Prcp_summer) %>% 
  left_join(Tmin_spring) %>% 
  left_join(Tmin_summer) %>% 
  left_join(Less10_var) %>% 
  left_join(GDU_spring) %>% 
  left_join(GDU_summer) %>% 
  rename(Place = place,
         Year = year) -> covariates

# Drop unused variables
rm(Prcp_spring, Prcp_var, Prcp_summer, Tmin_spring, Tmin_summer, Tmin_var, 
   GDU_summer, GDU_spring, Less10_var, daily_spring_summer)

```

Merge weather data with bat survey data and save it for further sessions

```{r}
# Merge data
df_clim <- df %>% 
  left_join(covariates)

# Export climate data for further sessions
save(df_clim, file = "bats_wClimate_2008-2019.Rdata")
```
